---
title: "Practice 4.1"
author: "Benson Shao"
date: "2024-10-20"
output: html_document
---
**Practice 4.1 Using the iris dataset, reproduce the plot below. Explore the function pairs allowing to generate a matrix of scatter plots between variables pairs (iris[1:4]). You will send me a pdf file of the plot your made. You can also add your code into a .Rmd in which you will import your image and knit it in a .html file. You will email me [vianneydenis@g.ntu.edu.tw] your files (and/or link) before next Monday so that I can review your work. The title of your email should be: `Practice 4.1 (your name: your student no.). ENJOY ;)**

Saving
```{r}
setwd("C:/Users/Esdeath/Desktop/testrepo")
pdf("iris_data.pdf", width = 8, height = 5)
```

Select data of each species
```{r}
setosa <- iris[iris$Species == "setosa" ,]
versicolor <- iris[iris$Species == "versicolor" ,]
virginica <- iris[iris$Species == "virginica" ,]
```

Draw the plot 
```{r}
#(1)Use "setosa" to draw the plot
plot(setosa$Petal.Length, setosa$Petal.Width,
     xlab = "Petal length (cm)", ylab = 'Petal width (cm)',
     xlim = range(iris$Petal.Length), ylim = range(iris$Petal.Width), 
     main = "Petal width vs. length")

#(2)Add the other species to the plot
points(versicolor$Petal.Length, versicolor$Petal.Width, col = 2, pch = 2)
points(virginica$Petal.Length, virginica$Petal.Width,  col = 3, pch = 3)

#(3)Add legends
legend("topleft",legend = c("setosa","versicolor","virginica"), col= 1:3,pch = c(1, 2, 3) ) 

#(4)Add the text and regression line
text(5, 0.5, "R= 0.96")
fit <- lm(Petal.Width ~ Petal.Length, data = iris)
abline(fit)

dev.off()
```

**Explore the function [pairs] allowing to generate a matrix of scatter plots between variables pairs (iris[1:4]).** 
```{r}
pairs(iris[1:4], 
      main = "Iris data matrix",
      col = iris$Species,  
      pch = 4)            
```



