---
title: "Practice 7.4"
author: "Benson Shao"
date: "2024-11-17"
output: html_document
---
## Practice 7.4 Create your own pearson.test function. (e.g. you will not use the above cor.test function but you will code your own function). The function should (at least) be able to test for the correlation between two variables. You will need to check at ?pt (or similar function) in order to extract the p-value. You can “dissect” the code source of the current cor.test function using getAnywhere("cor.test.default"). Using the rairuoho dataset, please use your customized function to test correlation between length at day 3 & day 4 then day 3 & day 8 in each treatment (nutrient and water). Interpret results.

```{r}
rairuoho<-read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt',header=T, sep="\t", dec=".", na.strings = "n/a")
```

## Define my pearson.test function
```{r}
pearson.test <- function(x, y) {
  n <- length(x)  # Sample size
  r <- cor(x, y)  # Pearson's correlation coefficient
  
  # Compute t-statistic
  t_stat <- r * sqrt((n - 2) / (1 - r^2))
  
  # Compute p-value (two-tailed test)
  p_value <- 2 * pt(-abs(t_stat), df = n - 2)
  
  # Return results
  return(list(correlation = r, t_statistic = t_stat, p_value = p_value))}
```

### Filter data by treatment
```{r}
nutrient_data <- subset(rairuoho, treatment == "nutrient")
water_data <- subset(rairuoho, treatment == "water")
```

### Test correlation: day3 & 4, day3 & 8 
```{r}
# Nutrient
nutrient_day3_day4 <- pearson.test(nutrient_data$day3, nutrient_data$day4)
nutrient_day3_day8 <- pearson.test(nutrient_data$day3, nutrient_data$day8) ### Not significant

# Water
water_day3_day4 <- pearson.test(water_data$day3, water_data$day4)
water_day3_day8 <- pearson.test(water_data$day3, water_data$day8)
```

## Create scatter plots with regression lines
```{r}
library(ggplot2)
```

### Define a plotting function
```{r, message=FALSE, warning=FALSE}
plot_regression <- function(data, x_var, y_var, treatment_filter, title) {
  # Filter data by treatment type
  filtered_data <- data[data$treatment == treatment_filter, ]

  ggplot(filtered_data, aes_string(x = x_var, y = y_var)) +
    geom_point(aes(color = treatment_filter), size = 3) +
    geom_smooth(method = "lm", se = TRUE, color = "black", linetype = "dashed") +
    labs(title = title, x = x_var, y = y_var) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "none")}

# nutrient: day3 & 4
plot1 <- plot_regression(rairuoho, "day3", "day4", "nutrient", "Nutrient: day3 & 4")

# nutrient: day3 & 8 (should not be plotted as not significant)
plot2 <- plot_regression(rairuoho, "day3", "day8", "nutrient", "Nutrient: day3 & 8")

# water: day3 & 4
plot3 <- plot_regression(rairuoho, "day3", "day4", "water", "Water: day3 & 4")

# water: day3 & 8
plot4 <- plot_regression(rairuoho, "day3", "day8", "water", "Water: day3 & 8")
```

### Draw all plots
```{r, message=FALSE, warning=FALSE}
library(gridExtra)
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2)
```
